Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-02-17 WSSSPE2_report/WSSSPE2_review.tex
Section headers: add visual aid.
Subsection headers: add visual aids.
Shorten URLs.

\section{Introduction}

@@

\section{Submissions, Peer-Review, and Peer-Grouping}

@@

\section{Keynotes}

\todo{VOLUNTEER NEEDED to write summaries for Kaitlin and Neil's keynotes}

\section{Defining Sustainability}

\todo{writing led by Dan - other contributions welcome}

@@

\section{Exploring Sustainability}

\todo{Colin to lead writing of this section}

Google doc notes - \url{https://docs.google.com/a/uchicago.edu/document/d/10XkshP3YjXnA5JQgP-UFHbOxshWczzehWGSo6V7bNH4/edit}

@@

\subsection{Papers}

\begin{itemize}
\item Mario {Rosado de Souza} and Robert Haines and Caroline Jay. Defining Sustainability through Developers' Eyes: Recommendations from an Interview Study~\cite{wssspe2_rosada_de_souza}
\item Robert Downs, W. Christopher Lenhardt, Erin Robinson, Ethan Davis, Nicholas Weber. Community Recommendations for Sustainable Scientific Software~\cite{wssspe2_downs}
\item Abani Patra, Matthew Jones, Steven Gallo, Kyle Marcus and Tevfik Kosar. Role of Online Platforms, Communications and Workflows in Developing Sustainable Software for Science Communities~\cite{wssspe2_patra}
\item Marlon Pierce, Suresh Marru and Chris Mattmann.{WSSSPE2}: Patching It Up, Pulling It Forward~\cite{wssspe2_pierce}
\item Justin Shi. Seeking the Principles of Sustainable Software Engineering~\cite{wssspe2_shi}
\item Colin C. Venters, Michael K. Griffiths, Violeta Holmes, Rupert R. Ward and David J. Cooke. The Nebuchadnezzar Effect: Dreaming of Sustainable Software through Sustainable Software Architectures~\cite{wssspe2_venters}
\end{itemize}

\subsection{Discussion and Actions}

\section{Software Development Experiences}

\todo{Nancy to lead writing of this section}

Google doc notes - \url{https://docs.google.com/a/uchicago.edu/document/d/1_q0jmiEPNFRGAEtGLk_QxQ1vVokgTDYruTGURowNGkY/edit}

@@

\subsection{Papers}

\begin{itemize}
\item Jordan Adams, Sai Nudurupati, Nicole Gasparini, Daniel Hobley, Eric Hutton, Gregory Tucker and Erkan Istanbulluoglu. Landlab: Sustainable Software Development in Practice ~\cite{wssspe2_adams}
\item Alice Allen and Judy Schmidt. Looking before leaping: Creating a software registry~\cite{wssspe2_allen}
\item Carl Boettiger, Ted Hart, Scott Chamberlain and Karthik Ram. Building software, building community: lessons from the {ROpenSci} project~\cite{wssspe2_boettiger}
\item Michael R. Crusoe and C.Titus Brown. Channeling community contributions to scientific software: a hackathon experience~\cite{wssspe2_crusoe}
\item Yolanda Gil,  Eunyoung Moon and James Howison. No Science Software is an Island: Collaborative Software Development Needs in Geosciences~\cite{wssspe2_gil}
\item Ted Habermann, Andrew Collette, Steve Vincena, Werner Benger, Jay Jay Billings, Matt Gerring, Konrad Hinsen, Pierre de Buyl, Mark K\"{o}nnecke, Filipe Rnc Maia and Suren Byna. The Hierarchical Data Format ({HDF}): A Foundation for Sustainable Data and Software~\cite{wssspe2_habermann}
\item Marcus Hanwell, Patrick O'Leary and Bob O'Bara. Sustainable Software Ecosystems: Software Engineers, Domain Scientists, and Engineers Collaborating for Science~\cite{wssspe2_hanwell}
\item Eric Hutton, Mark Piper, Irina Overeem, Albert Kettner and James Syvitski. Building Sustainable Software - The {CSDMS} Approach~\cite{wssspe2_hutton}
\item W. Christopher Lenhardt, Stanley Ahalt, Matt Jones, J. Aukema, S. Hampton, S. R. Hespanh, R. Idaszak and M. Schildhauer. {ISEES-WSSI} Lessons for Sustainable Science Software from an Early Career Training Institute on Open Science Synthesis~\cite{wssspe2_lenhardt}
\item Jory Schossau and Greg Wilson. Which Sustainable Software Practices Do Scientists Find Most Useful?~\cite{wssspe2_schossau}
\item James S. Spencer, Nicholas S. Blunt, William A. Vigor, Fionn D. Malone, W. M. C. Foulkes, James J. Shepherd and Alex J. W. Thom. The {H}ighly {A}ccurate {N-DE}terminant ({HANDE}) quantum {Monte Carlo} project: Open-source stochastic diagonalisation for quantum chemistry~\cite{wssspe2_spencer}
\end{itemize}

\subsection{Discussion and Actions}

\section{Credit \& Incentives}

\todo{Dan (OR VOLUNTEER) to lead writing of this section}

Google doc notes - \url{https://docs.google.com/document/d/1Jsi_trBT5cjEXVnGjGHJpcIhMEcVpAn6XmYNxfIHXv0/edit#heading=h.gk1dzxu9wi6q}

@@

\subsection{Papers}

\begin{itemize}
\item James Howison. Retract bit-rotten publications: Aligning incentives for sustaining scientific software~\cite{wssspe2_howison}
\item Daniel Katz and Arfon Smith. Implementing Transitive Credit with {JSON-LD}~\cite{wssspe2_katz}
\item Ian Kelley. Publish or perish: the credit deficit to making software and generating data~\cite{wssspe2_kelley}
\end{itemize}

\subsection{Discussion and Actions}

citation ecosystems sub-breakout: \url{https://docs.google.com/document/d/1SOja7k0mXwvOB2tQD3ZW3O38Qv4NwfA21zdMQ9JlUQc/edit}

Taxonomy of Contributorship/Guidelines for Software for Tenure Review sub-breakout: \url{https://docs.google.com/document/d/1sDZLaBR2elZ9Tj6eLIwEtf5VqQ_x_AN_SetKtJlFUpI/edit}

Prizes sub-breakout: \url{https://docs.google.com/document/d/1foSppeHQeQ-qOjd-Rp_CUCNy8JMLxd_vrCDMmqvG2-w/edit} (but as of 2/6, nothing is here and need to look at main breakout doc - Suresh is moving text, I think)

\section{Reproducibility \& Reuse \& Sharing}

\todo{Neil OR VOLUNTEER to lead writing of this section}

google doc notes - \url{https://docs.google.com/document/d/1Mivadj16_9tPrw8Nxmhp72AtNC0BEVO-f7M84FXTbSU/edit#heading=h.elaktnt2tep9}

@@

\subsection{Papers}

\begin{itemize}
\item Jakob Blomer, Dario Berzano, Predrag Buncic, Ioannis Charalampidis, Gerardo Ganis, George Lestaris and Ren\'{e} Meusel. The Need for a Versioned Data Analysis Software Environment~\cite{wssspe2_blomer}
\item Ryan Chamberlain and Jennifer Schommer. Using {Docker} to Support Reproducible Research~\cite{wssspe2_chamberlain}
\item Neil Chue Hong. Minimal information for reusable scientific software~\cite{wssspe2_chue_hong}
\item Tom Crick, Benjamin A. Hall and Samin Ishtiaq. ``Can I Implement Your Algorithm?'': A Model for Reproducible Research Software~\cite{wssspe2_crick}
\item Bryan Marker, Don Batory, Field G. Van Zee and Robert van de Geijn. Making Scientific Computing Libraries Forward Compatible~\cite{wssspe2_marker}
\item Stephen Piccolo. Building Portable Analytical Environments to improve sustainability of computational-analysis pipelines in the sciences~\cite{wssspe2_piccolo}
\end{itemize}

\subsection{Discussion and Actions}

\section{Code Testing \& Code Review}

\todo{Karen to lead writing of this section}

google notes doc - 
\url{https://docs.google.com/document/d/1G1_U5-hkAvDKtTgT4DvUhIYvyYUY4cBBvQHJIdgADa0/edit#heading=h.v7g0nqpn1fur}

@@

\subsection{Papers}

@@

\subsection{Discussion and Actions}

\section{Conclusions}

\todo{currently just pre-workshop stuff - needs to be updated}

@@

\appendix

\section{Attendees}
\label{sec:attendees}

The following is a partial list of workshop attendees who registered on the
collaborative notes document~\cite{WSSSPE2-google-notes} that was used
for shared note-taking at the meeting, or who participated in a breakout groups
and were noted in that group's notes.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Submissions, Peer-Review, and Peer-Grouping}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Keynotes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{VOLUNTEER NEEDED to write summaries for Kaitlin and Neil's keynotes}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Defining Sustainability}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{writing led by Dan - other contributions welcome}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Exploring Sustainability}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Colin to lead writing of this section.}

Google doc notes - \url{http://tinyurl.com/mpbhvyb}
%https://docs.google.com/a/uchicago.edu/document/d/10XkshP3YjXnA5JQgP-UFHbOxshWczzehWGSo6V7bNH4/edit

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{itemize}
\item Mario {Rosado de Souza} and Robert Haines and Caroline Jay. Defining Sustainability through Developers' Eyes: Recommendations from an Interview Study~\cite{wssspe2_rosada_de_souza}
\item Robert Downs, W. Christopher Lenhardt, Erin Robinson, Ethan Davis, Nicholas Weber. Community Recommendations for Sustainable Scientific Software~\cite{wssspe2_downs}
\item Abani Patra, Matthew Jones, Steven Gallo, Kyle Marcus and Tevfik Kosar. Role of Online Platforms, Communications and Workflows in Developing Sustainable Software for Science Communities~\cite{wssspe2_patra}
\item Marlon Pierce, Suresh Marru and Chris Mattmann.{WSSSPE2}: Patching It Up, Pulling It Forward~\cite{wssspe2_pierce}
\item Justin Shi. Seeking the Principles of Sustainable Software Engineering~\cite{wssspe2_shi}
\item Colin C. Venters, Michael K. Griffiths, Violeta Holmes, Rupert R. Ward and David J. Cooke. The Nebuchadnezzar Effect: Dreaming of Sustainable Software through Sustainable Software Architectures~\cite{wssspe2_venters}
\end{itemize}

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Discussion and Actions}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Software Development Experiences}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Nancy to lead writing of this section}

Google doc notes - \url{http://tinyurl.com/pn4eq8z}
%https://docs.google.com/a/uchicago.edu/document/d/1_q0jmiEPNFRGAEtGLk_QxQ1vVokgTDYruTGURowNGkY/edit

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{itemize}
\item Jordan Adams, Sai Nudurupati, Nicole Gasparini, Daniel Hobley, Eric Hutton, Gregory Tucker and Erkan Istanbulluoglu. Landlab: Sustainable Software Development in Practice ~\cite{wssspe2_adams}
\item Alice Allen and Judy Schmidt. Looking before leaping: Creating a software registry~\cite{wssspe2_allen}
\item Carl Boettiger, Ted Hart, Scott Chamberlain and Karthik Ram. Building software, building community: lessons from the {ROpenSci} project~\cite{wssspe2_boettiger}
\item Michael R. Crusoe and C.Titus Brown. Channeling community contributions to scientific software: a hackathon experience~\cite{wssspe2_crusoe}
\item Yolanda Gil,  Eunyoung Moon and James Howison. No Science Software is an Island: Collaborative Software Development Needs in Geosciences~\cite{wssspe2_gil}
\item Ted Habermann, Andrew Collette, Steve Vincena, Werner Benger, Jay Jay Billings, Matt Gerring, Konrad Hinsen, Pierre de Buyl, Mark K\"{o}nnecke, Filipe Rnc Maia and Suren Byna. The Hierarchical Data Format ({HDF}): A Foundation for Sustainable Data and Software~\cite{wssspe2_habermann}
\item Marcus Hanwell, Patrick O'Leary and Bob O'Bara. Sustainable Software Ecosystems: Software Engineers, Domain Scientists, and Engineers Collaborating for Science~\cite{wssspe2_hanwell}
\item Eric Hutton, Mark Piper, Irina Overeem, Albert Kettner and James Syvitski. Building Sustainable Software - The {CSDMS} Approach~\cite{wssspe2_hutton}
\item W. Christopher Lenhardt, Stanley Ahalt, Matt Jones, J. Aukema, S. Hampton, S. R. Hespanh, R. Idaszak and M. Schildhauer. {ISEES-WSSI} Lessons for Sustainable Science Software from an Early Career Training Institute on Open Science Synthesis~\cite{wssspe2_lenhardt}
\item Jory Schossau and Greg Wilson. Which Sustainable Software Practices Do Scientists Find Most Useful?~\cite{wssspe2_schossau}
\item James S. Spencer, Nicholas S. Blunt, William A. Vigor, Fionn D. Malone, W. M. C. Foulkes, James J. Shepherd and Alex J. W. Thom. The {H}ighly {A}ccurate {N-DE}terminant ({HANDE}) quantum {Monte Carlo} project: Open-source stochastic diagonalisation for quantum chemistry~\cite{wssspe2_spencer}
\end{itemize}

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Discussion and Actions}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Credit \& Incentives}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Dan (OR VOLUNTEER) to lead writing of this section}

Google doc notes - \url{http://tinyurl.com/k8ruyn9}
%https://docs.google.com/document/d/1Jsi_trBT5cjEXVnGjGHJpcIhMEcVpAn6XmYNxfIHXv0/edit#heading=h.gk1dzxu9wi6q

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{itemize}
\item James Howison. Retract bit-rotten publications: Aligning incentives for sustaining scientific software~\cite{wssspe2_howison}
\item Daniel Katz and Arfon Smith. Implementing Transitive Credit with {JSON-LD}~\cite{wssspe2_katz}
\item Ian Kelley. Publish or perish: the credit deficit to making software and generating data~\cite{wssspe2_kelley}
\end{itemize}

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Discussion and Actions}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
citation ecosystems sub-breakout: \url{http://tinyurl.com/l4tc6a4}
%https://docs.google.com/document/d/1SOja7k0mXwvOB2tQD3ZW3O38Qv4NwfA21zdMQ9JlUQc/edit

Taxonomy of Contributorship/Guidelines for Software for Tenure Review sub-breakout: \url{http://tinyurl.com/mx58tmx}
%https://docs.google.com/document/d/1sDZLaBR2elZ9Tj6eLIwEtf5VqQ_x_AN_SetKtJlFUpI/edit

Prizes sub-breakout: \url{http://tinyurl.com/mjok4o5}
%https://docs.google.com/document/d/1foSppeHQeQ-qOjd-Rp_CUCNy8JMLxd_vrCDMmqvG2-w/edit 
(but as of 2/6, nothing is here and need to look at main breakout doc - Suresh is moving text, I think)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Reproducibility \& Reuse \& Sharing}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Neil OR VOLUNTEER to lead writing of this section}

google doc notes - \url{http://tinyurl.com/kqpe87z}
%https://docs.google.com/document/d/1Mivadj16_9tPrw8Nxmhp72AtNC0BEVO-f7M84FXTbSU/edit#heading=h.elaktnt2tep9

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{itemize}
\item Jakob Blomer, Dario Berzano, Predrag Buncic, Ioannis Charalampidis, Gerardo Ganis, George Lestaris and Ren\'{e} Meusel. The Need for a Versioned Data Analysis Software Environment~\cite{wssspe2_blomer}
\item Ryan Chamberlain and Jennifer Schommer. Using {Docker} to Support Reproducible Research~\cite{wssspe2_chamberlain}
\item Neil Chue Hong. Minimal information for reusable scientific software~\cite{wssspe2_chue_hong}
\item Tom Crick, Benjamin A. Hall and Samin Ishtiaq. ``Can I Implement Your Algorithm?'': A Model for Reproducible Research Software~\cite{wssspe2_crick}
\item Bryan Marker, Don Batory, Field G. Van Zee and Robert van de Geijn. Making Scientific Computing Libraries Forward Compatible~\cite{wssspe2_marker}
\item Stephen Piccolo. Building Portable Analytical Environments to improve sustainability of computational-analysis pipelines in the sciences~\cite{wssspe2_piccolo}
\end{itemize}

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Discussion and Actions}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Code Testing \& Code Review}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{Karen to lead writing of this section}

google notes doc - 
\url{http://tinyurl.com/l5t5h45}
%https://docs.google.com/document/d/1G1_U5-hkAvDKtTgT4DvUhIYvyYUY4cBBvQHJIdgADa0/edit#heading=h.v7g0nqpn1fur

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Papers}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\subsection{Discussion and Actions}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{currently just pre-workshop stuff - needs to be updated}

@@

\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Attendees}  \label{sec:attendees}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The following is a partial list of workshop attendees who registered on the
collaborative notes document~\cite{WSSSPE2-google-notes} that was used
for shared note-taking at the meeting, or who participated in a breakout groups
and were noted in that group's notes.
